---
title: "Lab 10 Halloween Analysis"
author: "Shane Desfor"
format: html
---

## Lab 10 Halloween Data analysis

```{r}
candy_file <- "candy-data.csv"
##Set Row Names to First Column:

candy = read.csv("candy-data.csv", row.names=1)
sum(candy$fruity)

```
>Q1. How many different candy types are in this dataset? Ans. nrow(candy), 85
>Q2. How many fruity candy types are in the dataset? Ans., 38
The functions dim(), nrow(), table() and sum() may be useful for answering the first 2 questions.

```{r}
nrow(candy)
sum(candy$fruity)

```




>Q3. What is your favorite candy in the dataset and what is it’s winpercent value? 
Can Call row.names to see all the candies. Then call win percent of that candy. Ans.84%
>Q4. What is the winpercent value for “Kit Kat”? Ans. 77%
>Q5. What is the winpercent value for “Tootsie Roll Snack Bars”? Ans. 50%

```{r}
row.names(candy)
candy[ "ReeseÕs Peanut Butter cup",  ]
candy [ "Kit Kat", ]
candy[ "Tootsie Roll Snack Bars",] 

```


#Can pull a function from a package instead of moving through library

>Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset? Winpercent
>Q7. What do you think a zero and one represent for the candy$chocolate column? Answ. Percent of chocolate candy

```{r}

skimr::skim(candy)


```





>Q8. Plot a histogram of winpercent values
>Q9. Is the distribution of winpercent values symmetrical? Ans. No
>Q10. Is the center of the distribution above or below 50%? Ans. Below



```{r}
library("ggplot2")

ggplot(candy, aes(winpercent, bine=10,)) + geom_histogram()

```




```{r}
head(candy)
```



>Q11. On average is chocolate candy higher or lower ranked than fruit candy? chocolate is 61%, fruity is 44%
>Q12. Is this difference statistically significant? Yes, p-value = 2.871e-08

```{r}
chocolate.ind <- as.logical(candy$chocolate)
candy[ chocolate.ind, ]

##This is how you call winpercent OF chocolate logical

chocolate.win <- candy[ chocolate.ind, ]$winpercent

##Then to summarize, call mean
mean(chocolate.win)


##Now fruity
fruity.ind <- as.logical((candy$fruity))
fruity.win <- candy[fruity.ind, ]$winpercent
mean(fruity.win)

t.test(fruity.win, chocolate.win)


```


##3. Overal Candy Rankings
can use the sort function and order function 

```{r}
x <- c(5, 6, 1, 9, 10)
sort(x)
  
```

order shows the order of the vector, where each value shows the place the element should be in that order
but put it as a vector of location and order, get them ordered.
```{r}
order(x)
x [ order(x)]
```

>Q13. What are the five least liked candy types in this set? Nik Lnip, Boston Baked Beans, Chiclets, Super Bubble, Jaw Busters
>Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
order(candy$winpercent)
candy [order(candy$winpercent), ]

head(candy[order(candy$winpercent),], n=5)

#Tidy vers version
library("dplyr")
candy %>% arrange(winpercent) %>% head(5)

```


>Q15. Make a first barplot of candy ranking based on winpercent values.
Use geom col so all columns are an individual variable and plot

```{r}


ggplot2:: ggplot(candy, aes(x=winpercent, y=rownames(candy))) + geom_col()


```

Lets ReORder these

```{r}
ggplot(candy, aes(x=winpercent, y=(reorder(rownames(candy),winpercent)))) + geom_col()
```
Lets set a color vector (by candy type) that we can 
Color for every bar by candy type
nwor makes a black for every row in candy df

```{r}
my_cols <- rep("black", nrow(candy))
#Makes all rows black
#Then we say, for every TRUE this is a chocolate, then color it chocolate
my_cols[ as.logical(candy$chocolate)] <- "chocolate"
my_cols[ as.logical(candy$bars)] <- "brown"
my_cols[ as.logical(candy$chocolate & candy$bar)] <-"blue"
my_cols[ as.logical(candy$fruity)] <- "red"

               
```

> Q16, order them by winpercent and graph
> Q17. What is the worst ranked chocolate candy? Sixlets
> Q18. What is the best ranked fruity candy? Starburst

```{r}
ggplot(candy, aes(x=winpercent, y=(reorder(rownames(candy),winpercent)))) + geom_col(fill=my_cols)
```

#Looking at pricepercent
value for money?
what is the best candy for the least money
a way to do that is 'winpercent' v. pricepoercent
```{r}
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + geom_text(col=my_cols)

```
Make these points not overlap as much with gg repel
```{r}
library(ggrepel)

ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) +  geom_text_repel(col=my_cols, size=2.0, max.overlaps = 5)



```


>Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck? Recess miniatures
Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular? Nik L Nip, Nestle Smarties, Ring pop, Hersheys Krakel, and Hersheys chocolate. Least popular is Nik L Nip

```{r}
ord_win <- order(candy$winpercent, decreasing = TRUE)
head(candy [ord_win, c(11,12)])

ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )

```



##Sec. 5, exploring coorelation 
>Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?
>Q23. Similarly, what two variables are most positively correlated?

```{r}
library(corrplot)
## corrplot 0.90 loaded
cij <- cor(candy)
corrplot(cij)
```


##PCA 
>Dont forget to set scale to true, its set to scale = false naturally, so that it is scaling varience instead of drastic values, like win percent.

```{r}
pca <- 
prcomp(candy, scale=TRUE)
summary(pca)
```


lets plot data on these new axis now. ie pc1 v pc2
```{r}
plot(pca$x[ , 1], pca$x[ , 2], col=my_cols, pch=16)


```

note, gg plot needs a data frame, so need to make a data frame of pca. So we bind the PCA data and add 3 new columns to the cady data frame
```{r}
# Make a new data-frame with our PCA results and candy data
my_data <- cbind(candy, pca$x[,1:3])
```


```{r}
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols, ) + geom_text_repel(max.overlaps = 5)

p
```



Can also show how much EACH GROUP, each variable is contributing to the variation making up the PCA axis. This is what the rotation readout is.
>Q24.What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?
Hard, fruity, and comes in multiples seems to pull the variation the most in the positive direction, making up the positive axis of the PC1 axis of the PCA plot.

```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

