---
title: "COVID-19 Variant Analysis"
author: "Shane Desfor - PID: A59018963"
format: html
---

## Step-By-Step
#Read in and examine csv file.
```{r}

covid <- read.csv("covid19_variants.csv")
head(covid)

```

#Going to need ggplot for this.
```{r}
library(ggplot2)
```

#...Realized waaay down below that this needed to change. And it needed to be here because of the way I had layered in date.
```{r}
covid$date <- as.Date(covid$date, format = "%Y-%m-%d")
```


#After much tweeking, I got a workerable plot to play with, but it needs some tuning in.
```{r}
covid_plot <- ggplot(covid, 
                     aes(x=date, 
                         y=percentage, 
                         color=variant_name))
covid_plot + 
  geom_point() + 
  labs(title = "COVID-19 Variants in California", 
       y= "Percentage of Sequenced Specimens", 
       color="Variant")
```


#First, remove the 'Total' from being plotted
```{r}
covid_filtered <- covid[covid$variant_name != "Total", ]
```


#Could not get geom_line to work, would not display any variants, going to try and group by variant, as a factor.
```{r}
covid_filtered$variant_name <- factor(covid_filtered$variant_name)
```

#Adjusted line plot
```{r}
covid_plot2 <- ggplot(covid_filtered, 
                      aes(x = date, 
                          y = percentage, 
                          color = variant_name, 
                          group = variant_name)) + 
  geom_line()
covid_plot2

```
#Worked! But it's a little sloppy.

#Added lables and adjusted the x-axis...still needs some more tweeking though.
```{r}
covid_plot2 + 
  labs(title = "COVID-19 Variants in California", 
       x= NULL, 
       y= "Percentage of Sequenced Specimens", 
       color="Variant") + 
  scale_x_date(date_labels = "%B %Y", date_breaks = "1 month") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#Get better x-axis starting and ending points, also discovered I needed a 'breaks vector'
```{r}
start_date <- min(covid_filtered$date)
end_date <- max(covid_filtered$date)
breaks.vec <- seq(min(covid_filtered$date), 
                  max(covid_filtered$date), by= "1 month")
```


##Final Product
#This is the one!
```{r}
covid_plot2 + 
  labs(title = "COVID-19 Variants in California", 
       x= NULL, y= "Percentage of Sequenced Specimens", 
       color="Variant") +
  scale_x_date(breaks = breaks.vec, date_labels = "%B %y") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```














